%------------------------------------------------------------------------------
%
%	money.tex General monetary system part
%
%	INCLUDE FILE FOR LaTeX2e DOCUMENT
%
%	AUTHOR: Ari Potkonen /JARVENPAA/ Mon Jun 28 2022
%------------------------------------------------------------------------------
%         1         2         3         4         5         6         7
%123456789012345678901234567890123456789012345678901234567890123456789012345678
%-BEGIN OF INCLUDE FILE--------------------------------------------------------
\part{Monetary system}
\chapter{Monetary system}\label{monetary_system}\index{monetary system}
Monetary system development and extension need. Monetary system what national
and union banks maintain is left behind from native digital citizens and
market digital development needs. It's time to tackle digital trade problems
by rising abstraction level of exchanged valuables and same time removing
trust problems with the frame agreement defining standardized digitalized and
supported business models on agreement and trade platform level. And maybe
later on we have laws supporting frame agreement based market operation. To
really to do frictionless trade we have to digitalize and standardize goods
and services digital presence at least on ecosystem frame agreement level.
What is product or service digital presence - asset - sold on digital market
before any products or services are delivered? Or does the vendor delivery
promise - certificate - be the thing to sell?

I would like to say that in digital trade, e-commerce, - asset - is product
marketing documentation offering, customer need fulfilling solution, with
customer value bigger that price value for customer, and sale is happening if
products are available, meaning if vendor given - certificate - "promise" from
availabi\-li\-ty is available and delivery terms are matching to customer's
customer value idea customer has got from product documentation called asset.
So in this document: Asset is the product or service description. Certificate
is asset defined product or service delivery promise to certain place at
certain time, at certain price. Certificate (vendor-promise) is what is
actually sold! And we need to be able to do sales transactions electrically.

\section{Traditional}
Traditional monetary system. Legal system is set so that generally banks are
only legal entities legitimized to produce state backed money, money which has
legally supported status and known behaving model. Amount of money on market
affects to it's value on market. If amount of money is small compared to
market size then money desirability rise because everyone on market want to
change they goods and services. This limits trading and rise of product price,
because most to offer for exchange medium wins medium and does exchange. If
others still exchange they products against smaller amount of money then money
value is grown and that is called to deflation. If there are lot of money on
market then all can do exchange at the same time without limitations and then
asked money amounts from same products grow because there is lot of money
available and can be asked more. This is called inflation which means change
medium value decrease. To control market operation governments, national and
unions central banks have some operations to manage total amount of money on
market as well control inflation or deflation on market.
% Faults Leaks Cumulation Fixes Extension
\newline

Increasing money amount on market:
\begin{itemize}
\item Release tangible money to circulation
\item Release intangible money to circulation
\item Negative Interest of loan
\item Positive Interest of saving
\item Loan of money
\item Purchase of bonds by central banks
%\item Government supported, financed things
%Relativistic Local/Global Zero Sum game operations are not listed:
%\item Local current revaluation (indirectly devaluates other currencies
%if/when there is any trade)
\end{itemize}

Decreasing money amount on market:
\begin{itemize}
\item Returned tangible money from circulation
\item Returned intangible money from circulation
\item Positive interest of loan
\item Negative interest of saving
\item Payment of loaned money and positive interest
\item Non-returned worn or lost tangible money
\item Non-returned corrupted or lost intangible money
%\item Government tax
%\item Government toll, decrease global money amount, when government money
%equated to central bank money
%Relativistic Local/Global Zero Sum game operations are not listed:
%\item Local current devaluation (indirectly revaluates other currencies
%if/when there is any trade)
\end{itemize}

\section{Electric money}
There are several intangible electric money activities ongoing.

Most of electric money systems are based to blockchains and cost of proof,
which due distributed nature is high enough to protect authenticity of each
transaction. Even this cost is not shown for the actual user, because there
is inbuild mechanism to finance cost by adding more money to circulation, it
still has high cost. This is also reason why we should not use these methods,
because cost will be taken from society and environment we live in form of
high electricity consumption or in form of exessive storage media material
consumption.

Blockchain based payment systems also try to keep anonymization by using
account or wallet numbers instead of actual owner names. Therefore systems have
been used a lot for the shadow transactions not tolerating full day light.

It might be nice to have nameless system for the device having credit used to
purchase goods and services, but does it bring any significant benefits is open
question.

It's hard to say do we really get any official legally supported electric money
which removes possibility to have full payment chain visibility or possibility
to do full payment chain resolution afterwards.

Payment chain visibility make possible to eliminate fraud from market by
removing cheating player by allowing to detect cheating member.

There are
setups which are called to e-money even those are actually electric
transactions between service provider maintained bank accounts.

It's possible to automate old check system, by doing full signature and
encryption chain where each owner does check signature and encryption with
receiver public key. But then it's obvious who had participated into payment
chain.

Pure electric bank note initiated and signed by your bank, or central bank,
against money on your account is easily under thread if first receptive is not
the last one too. Check note is encrypted with your public encrypting key for
delivery. When used it's send to receiver with public key encryption. During
transfer everyone store instance encryption is removed and new done by each
holder. Final receiver is again a bank and it pays amount of money for sender
against the note. Final receiver name is stored by bank to allow fraud
detection, by backward tracing. With successful access chain all mid-holders
are not known. Every owner knows and stores last and next legitime owner, but
not the whole chain.

Because these security problems direct bank transfers are preferred on
electric trade.

\section{Quantitative Easing}
Era of Quantitative Easing (QE). It has been thought that peoples and
organizations who did and do most from new productive investments finance
those investments with loan and therefore giving them benefit with negative
interest rates and junk bond purchases and acceptance to mortgage gives best
boost to economy. But that is not true anymore, not even been when
quantitative easing started at 2008 crisis. Those organizations had put easy
money to living and unproductive investments without actually investing to
sustainable growth. More and more money goes to easy wins and protecting they
position and status on earth, our market place. Protecting they own living
standard without any even small willingness to change sustainable business
models. Why? They think that it's easier to protect existing just now than
create new business. And in short time it be so, allowing to transfer more
money to consumption; cars, houses, cottages, jets, boats, second, third,
fourth houses, villas with personnel and maintenance, even if not used at all.
So most usable money went to consumption, not to sensible new business
creation nor social investments like education etc. Even central bank has
supported this old, mostly by fossil fuels driven economy, consumption style
which is coming to dead end because our market place earth doesn't tolerate
it. There isn't anymore need to blow up climate satellite with normal bolt set
between rocket phases where you should have exploding one. Everyone here can
notice change even without that satellite. Glaciers are melting and drought
follows here and there. And those bosses used most of money to they own
consumption are now coming to dead end, some even started wars to keep
existing fossil backed business model and gain more for themselves. It does
not work anymore, everyone will loose if continuing that way, we are not sure
can we get this thing turned right even if we all work seamlessly together
towards sustainable goal where economy expansion is on space and sustainable
business models on earth. Extinction is on offer plate if we doesn't
understand common good. Money wise -- because money is only number, growth
can continue on earth, but is more or less social services growth, recycling
growth and technology improvement. So, we should be more on quality growth,
than volume growth side, if we want to survive.

It's is good to notice that all entrepreneurs do not belong to this greedy
group and have grown they business mostly with income financing and caring
from sustainability. Still amount of outdated business and nonproductive play
with financial market without actually producing any good is taking too big
part from whole economy. Therefore artificial respiration of these dead
businesses with QE is coming to end. These not-trustable, cheating,
frightening and even violent organizations has to be let to stay on they own
and concentrate to original producer or service creator possibility to do
business straightly with the customer. Then environment aware peoples can vote
with they wallet when system has been built right. Business natural tendency
to renew and prune dead ends has to be restored by ending QE. QE successor
(Sustainable Financing \cite{EU_Sustainable_finance}) has to have long term
sustainability or space utilization targets for expansion. Dead end public
support financing has to end. Negative interest rate is normally taken from
normal worker's back by reducing the value of his/her money, and giving it to
those who has too much already, and whose are already underperforming by
sticking to old world, grabbing only very short term local wins, and spoiling
larger community possibilities to live good life. Risk is that autocratic
chieftains drive they own very short term benefits and peoples resisting the
change are supporting these chieftains against they own and they children
benefit. Then the result is global repetition from what happened on Easter
Island on earth last resort where we get clan chieftain moʻai repetition when
all key resources are used and system collapses.

\section{Trust}\label{trust}\index{trust}
Legal basement for monetary system requires definition of legal entities
definitions, then actual law text could refer to these defined legal entities
and define they roles, rules and consequences of non-compliance. Most used
legal entity roles are; natural person, corporate body, commune, bank, nation,
national property register, national bank, national tax, national toll,
central bank, central tax, central toll, exchange and national archive.
Operations these several legal entities can do are for example: exchange (sell
\& buy), tax, toll, loan, mortgage, refinance (national and central banks),
rate (national and central banks), etc. Exchange of goods and services using
money and money kind abstract valuables is based to peoples trust. Governments
have based legal framework to maintain peoples trust to these valuables to
make goods and services exchange easy and effective.

\section{Exchange}\label{exhange}\index{exhange}
When peoples do something they need others help to reach bigger advancements
that they can do alone. To be fear there is need for reciprocity and it's hard
to archive in bigger scale without some medium of exchange. Therefore some
selected valuable easy to exchange commodities started to be used as medium of
exchange. While economy grow more there were greater need to have more
abstract medium of exchange and legal basement for that created to maintain
peoples trust to it. Exchange valuables; coin, bank note, dept, bond,
commodity, share, property, mortgage, derivative and certificate. This
certificate and operations for it are the new things we are interested to
define for digital trade, e-commerce.

%\begin{figure}
% \begin{center}
%  \epsfig{figure=money.eps,width=1.515in,height=4.895in}
%  \caption{Monetary system}
%  \label{fig:MonetarySystem}
%\end{center}
%\end{figure}

%-END OF INCLUDE FILE----------------------------------------------------------

