%------------------------------------------------------------------------------
%
%	preface.tex Document preface part
%
%	INCLUDE FILE FOR LaTeX2e DOCUMENT
%
%	AUTHOR: Ari Potkonen /JARVENPAA/ Mon Jun 28 2022
%------------------------------------------------------------------------------
%         1         2         3         4         5         6         7
%123456789012345678901234567890123456789012345678901234567890123456789012345678
%-BEGIN OF INCLUDE FILE--------------------------------------------------------
PREFACE
\label{preface}
\index{preface}
\vskip\baselineskip
\addcontentsline{toc}{chapter}{Preface}
As part of my Digital Commerce\cite{DigitaalinenKauppa} studies I decided to do vision document from electric-commerce, trading by collecting early 21st century ecosystem ideas together. Starting from traditional GSM barefoot data delivery network using charge stations as distribution centers on rural area, continuing to ideas from Debian ".dep" and Fedora ".rpm" package systems generalization with XML. Created XML blocks stored to EU digital preservation archive block store and from there lifted to cloud services object store at cloud edge for tradesystem use. Services dynamically scaled at cloud edge on need bases. On cloud edge asset defining XML is holding or pointing to product official documentation, what to purchase, and other XML defining how product or service is available at time of purchase for location of purchase. Then trade transaction support is created to sale these latter XML documents. This is opened more in document if not clear now ;-). Latest ideas collected from digital commerce training, related books and from Nordic Retail and Wholesale Association (NRWA) Conference (NRWC) at Tampere UNIveresity (TUNI) 2022 November. Extending, merging and combinint those together here, and you are welcomed to expand our understanding. Please feel free to utilize, comment or improve these documents, this and appendixes included.  I came to this openness and publicity decision because digital commerce ecosystem creation is community action which require wide cooperation to protect and optimize every party position in community while respect the others. So it's a compromise for all, but meant to be more economic and effective too than original situation at the beginning. National governments usually participate more afterwards to work, but EU is active and there is risk that EU (Cyber Resilience Act (CRA)\cite{EU_CRA}) spoils possibility to have good ecosystem by demanding cryptography use without implementing open reference implementations first, nor needed legal framwork either. There is need for open hardware and open software reference implementations before demanding it to be used. It is EU government's disease that it's so much easier to restrict and demand that actually do, finance and supervise practical implementation of something useful needed here. Just demanding and ruling leads to monopolistic solution, very significantly limiting citizens and small business rights and possibility innovate, build they business ahead. They do not understand that for every software they required or allowed to be encrypted for trade they has to have digital preservation archive instance for unencrypted originals, like for books have free copy right for national archives to get some originals saved for forthcoming generations. Same has to happen in this new digital ecosystem to be fair from generations to generations. That has to be done first, and lawmakers clearly don't understand it now. Archives are created and used for documents, library and museum data, but not for pure trading. In future traded digital originals, the only originals, has to be archived before trading. Trading instances, encrypted or not, are lifted from archive services to cloud edge for trade on need bases. Asset -- documentation is there all time, blobs are lifted based on archive servers load. That's it, more details on few next pages. Hopefully you get some ideas from here for your professional life!
%\vskip\baselineskip
Sincerely yours Ari Potkonen
%\vspace*{\fill}
%\cite{ISO_IEC_19790}, 
%-END OF INCLUDE FILE----------------------------------------------------------
